---
title: "Practical Machine Learning Writeup"
author: "jpesic"
date: "March 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background



Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict if the barbell lifts are performed correctly. 
The subjects were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

# Data Import

```{r data import, message=FALSE}
require(readr)
setwd("~/R Scripts/R for Data Science/Machine Learning/Final Project")

pml_training <- read_csv("pml-training.csv")
pml_testing <- read_csv("pml-testing.csv")
```

# Data Cleaning

Columns with missing inputs are identified; a new data frame without those columns is generate. In addition, columns without general predictive values (subject names/time stamps) are deleted from the training data frame. "Classe" column is converted to factor format.

```{r data clean}
columns<-sapply(pml_training, function(x) !any(is.na(x)))
pml_training_na<-pml_training[,colnames(pml_training)%in%names(which(columns==TRUE))]
training<-pml_training_na[,8:57]
training$classe<-as.factor(training$classe)
```


## Modeling 

Thusly generated "training" dataframe is split into two data.frames, one for trainig purposes and another for validation. "RandomForest" package is used to model the data.  

```{r model}
require(caret)
require(randomForest)
set.seed(123)
#splt orignal training model into two
inTrain <- createDataPartition(training$classe, p = 0.70, list = FALSE)
validation <- training[-inTrain, ]
training_new <- training[inTrain, ]

# try train (random Forest)
model1<-randomForest(classe~., data=training_new)

```

## Validation

"validation" data frame is used to validate Random Forest model. The accuracy of the model is  calculated from the number of the off-diagonal (misidentified) elements. The accuracy is >0.993.
```{r validate, echo=TRUE}

validate1<-predict(model1, validation)

table(validate1, validation$classe)


```

## Final Quiz (Testing) Data Prediction

Testing data frame is processed to include the same columns as used to train the model. Then, the final prediction is generated.

```{r test, echo=TRUE}

param<-colnames(training_new)

testing<-pml_testing[,colnames(pml_testing)%in%param]

head(testing)

quiz_prediction<-predict(model1, testing)

print(quiz_prediction)


```






